---
title: "week_eight_lee_d"
author: "Dong Lee"
date: "2025-05-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Research question

A researcher hypothesizes that kids watching TV will result in an increase in the number of pieces of candy corn eaten. The kids are subjected to 5 hours of television. The number of candy corn candies consumed during a 10-minute period is counted both with and without TV.

### Data

```{r data}
# Create a data frame for the candy corn consumption data
candy_data <- data.frame(
  Experiment = 1:10,
  With_TV = c(12, 7, 3, 11, 8, 5, 14, 7, 9, 10),
  Without_TV = c(8, 7, 4, 14, 6, 7, 12, 5, 5, 8)
)

```

### Computation

``` {r compute}
#Calculate the differences (With TV - Without TV)
candy_data$Difference <- candy_data$With_TV - candy_data$Without_TV

#mean of differences
mean_diff <- mean(candy_data$Difference)

#variance and standard deviation of differences
var_diff <- var(candy_data$Difference)
sd_diff <- sd(candy_data$Difference)

cat("Mean =", mean_diff, "\n")
cat("S =", sd_diff, "\n")
cat("S² =", var_diff, "\n")
```

The mean difference of 1 represents that the average increase in candy corn consumption when kids watch TV compared to when they don't. A positive value of 1 indicates that kids eat more candy corn when watching TV on average.

The standard deviation of 2.404 represents spread of the differences between observations. It indicates how consistent watching TV affects candy corn assumption across kids. A larger S suggests more variability.

The variance of 5.778 represents another measure of variability in the differences. It is important for calculating standard error in the t-test.

### T-test analysis

```{r t-test}
# calculate standard error
n <- nrow(candy_data)
se <- sd_diff / sqrt(n)

# t-statistic
t_stat <- mean_diff / se

# degrees of freedom
df <- n - 1

# Calculate p-value (two-tailed test)
p_value <- 2 * pt(abs(t_stat), df, lower.tail = FALSE)


cat("t-statistic =", t_stat, "\n")
cat("Degrees of freedom =", df, "\n")
cat("p-value =", p_value, "\n")

# verifying with t.test function
t_test_result <- t.test(candy_data$With_TV, candy_data$Without_TV, paired = TRUE)

t_test_result
```

The computed t-statistic is 1.316. This value represents how many standard errors the sample mean difference is away from the null hypothesis value.

The null hypothesis in this study is that there is no difference in candy corn consumption between the "With TV" and "Without TV" conditions.

$H_0: \mu_d = 0$ where $\mu_d$ represents the true mean difference in the population.

The alternative hypothesis is that there is a difference in candy corn consumption between the "With TV" and "Without TV" conditions.

I chose the conventional probability level of 0.05 for this study. This means that we would rejject the null hypothesis if there is less than a 5% chance of observing the data if the null hypothesis were true. This value balances the risks of Type I error of falsely rejecting a true null hypothesis and Type II error of failing to reject a false null hypothesis. And this level is also a standard practice in social science research.

The degrees of freedom for a paired t-test is n-1, where n is the number of pairs. In this study, degrees of freedom = 10 - 1 = 9.

With a t-test statistic of 1.316 and 9 degrees of freedom, the p-value is approximately 0.22 for two-tailed. Since the p-value is greater than the alpha value of 0.05, we fail to reject the null hypothesis. This means that although the kids consumed on average 1 more candy corn when watching TV versus not watching TV, this difference is not statistically significant. The observed difference could be due to random chance rather than a true effect of watching TV on candy corn consumption.







